---
title: "summary-results-plots"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r}
#| echo: false
#| warning: false
#| message: false
#run this and then everything is ready to go
library(here)
source(here("scripts","scripts-heat-hosp-green-hia", "analysis-summary-attrib-hosps.R"))
```


# Scenario definitions

-   Tree canopy

    -   Population-based Scenario: AI: Increase by 10% in all zip codes

    -   Targeted

        -   Scenario AII1: Increase by 10% in zip codes in the lowest 1/5th of current TC cover (i.e. \<=20th pctile)

        -   Scenario AII2: Increase by 10% in zip codes in the highest 1/5th of the Social Vulnerability Index (i.e. \>80th pctile)

        -   Scenario AII3: Increase by 10% in zip codes in the highest 1/5th of hospitalization burden (i.e. \>80th pctile)

    -   Proportionate-universalism

        -   Scenario AIII1: Increase by 10% for bottom 1/5th of current TC cover... down to 2% for top 1/5th

        -   Scenario AIII2: Increase by 10% for top 1/5th of SVI ... down to 2% for bottom 1/5th

        -   Scenario AIII3: Increase by 10% for top 1/5th of hospitalization burden ... down to 2% for bottom 1/5th

-   Impervious surface cover

    -   Population-based: Scenario BI: Decrease by 10% in all zip codes

    -   Targeted

        -   Scenario BII1: Decrease by 10% in zip codes in the highest 1/5th of current imperv cover (i.e. \>80th pctile)

        -   Scenario BII2: Decrease by 10% in zip codes in the highest 1/5th of the Social Vulnerability Index (i.e. \>80th pctile)

        -   Scenario BII3: Decrease by 10% in zip codes in the highest 1/5th of hospitalization burden (i.e. \>80th pctile)

    -   Proportionate-universalism

        -   Scenario BIII1: Decrease by 10% for top 1/5th of current imperv cover ... down to 2% for bottom 1/5th

        -   Scenario BIII2: Decrease by 10% for top 1/5th of SVI ... down to 2% for bottom 1/5th

        -   Scenario BIII3: Decrease by 10% for top 1/5th of hospitalization burden ... down to 2% for bottom 1/5th
        
# Summary results (Table 2)
## Keeping denominator the same across scenarios

```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop=n_cases_diff_wide_by_scenario %>% 
  left_join(attrib_hosps_for_summary_nogeo,by="zcta") %>% 
#  filter(zcta_intervened==1) %>% #limit to intervention zip codes
  group_by(scenario,scenario_intervention,scenario_type_3,scenario_type_7) %>% 
  summarise(
    n_zcta=n(),
    pop=sum(pop,na.rm=T),
    attrib_hosps=sum(attrib_hosps,na.rm=T),
    n_cases_diff=sum(n_cases_diff,na.rm=T)
  ) %>% 
  mutate(
    ir_baseline=attrib_hosps/pop,
    ir_alt=(attrib_hosps-n_cases_diff)/pop,
    ird=ir_alt-ir_baseline,
    irr=ir_alt/ir_baseline
  ) %>% 
  ungroup()

attrib_hosps_by_scenario_same_pop %>% 
  dplyr::select(-contains("scenario_type_3")) %>% 
  knitr::kable(
    digits=c(0,0,0,
             0,0,
             1,3,
             9,9,9,3)
  )
```


## Restricting denominator to those zip codes intervened upon
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop=n_cases_diff_wide_by_scenario %>% 
  left_join(attrib_hosps_for_summary_nogeo,by="zcta") %>% 
   filter(zcta_intervened==1) %>% #limit to intervention zip codes
  group_by(scenario,scenario_intervention,scenario_type_3,scenario_type_7) %>% 
  summarise(
    n_zcta=n(),
    pop=sum(pop,na.rm=T),
    attrib_hosps=sum(attrib_hosps,na.rm=T),
    n_cases_diff=sum(n_cases_diff,na.rm=T)
  ) %>% 
  mutate(
    ir_baseline=attrib_hosps/pop,
    ir_alt=(attrib_hosps-n_cases_diff)/pop,
    ird=ir_alt-ir_baseline,
    irr=ir_alt/ir_baseline
  )%>% 
  ungroup()

attrib_hosps_by_scenario_restrict_pop %>% 
  dplyr::select(-contains("scenario_type_3")) %>% 
  knitr::kable(
    digits=c(0,0,0,
             0,0,
             1,3,
             9,9,9,3)
  )
```

## Boxplots
Summary of summary results over scenarios
### No change to denominator

#### Difference in number of cases (absolute benefit)

Summarized by intervention targeting type.
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=n_cases_diff))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp).
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=n_cases_diff))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```

#### IRD
Summarized by intervention targeting type
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=ird))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp)
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=ird))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```

#### IRR
Summarized by intervention targeting type.
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=irr))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp).
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_same_pop %>% 
  ggplot(aes(y=irr))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```


### Restricting denominator

#### Difference in number of cases (absolute benefit)

Summarized by intervention targeting type.
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=n_cases_diff))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp).
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=n_cases_diff))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```

#### IRD
Summarized by intervention targeting type.
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=ird))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp).
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=ird))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```

#### IRR
Summarized by intervention targeting type.
Both tree canopy and impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=irr))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
#    rows=vars(scenario_intervention),
    cols=vars(scenario_type_3)
  )
```

Summarized by intervention type (tree vs imp).
```{r}
#| echo: false
#| warning: false
#| message: false
attrib_hosps_by_scenario_restrict_pop %>% 
  ggplot(aes(y=irr))+
  geom_boxplot()+
  theme_bw()+
  facet_grid(
        cols=vars(scenario_intervention)
  )
```
