---
title: "Interactive maps: estimating benefits of greening on heat-related hospitalizations"
author: "Michael D Garber"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r}
#| echo: false
#| warning: false
#| message: false
#Load data 
#March 13, 2025
#copied from climate-health/docs/heat-hosp-green-hia/vis-interactive-maps.qmd
#as tmap interactive and quarto don't seem to work well right now
#Updated with tmap v 4 code as well
library(here)
library(mapview)  
library(tidyverse)
library(viridis)
library(viridisLite)
library(leaflet)
library(leaflet.extras)
library(leafgl)
library(cols4all)
library(tmap)
library(tmaptools)
library(RColorBrewer)
source(
  here("scripts","scripts-heat-hosp-green-hia","read-attrib-hosps-ratios-diffs.R"))
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")
load("attrib_hosps_for_summary_nogeo.RData")
```


```{r}
#| echo: false
#| warning: false
#| message: false
#define some color palettes that flip at 1
#Mar 13, 2025: I don't use these palettes in tmap
pal_rdylgn5=RColorBrewer::brewer.pal(n=5,"RdYlGn")
pal_rdylgn5_rev=rev(pal_rdylgn5)

pal_rdylbu=RColorBrewer::brewer.pal(n=5,"RdYlBu")
pal_rdylbu_rev=rev(pal_rdylbu)

viridis_5=viridis(5)
```

# Baseline maps: hospitalizations and covariates
## Attributable number of hospitalizations per heat-wave day (baseline)

```{r}
#| echo: false
#| warning: false
#| message: false
tmap_mode("view") #run again just in case it was changed above.

tmap_attrib_hosps= zcta_ca_geo_simplified %>% 
  dplyr::select(-contains("pop")) %>% 
  dplyr::select(-contains("zcta_in_cali")) %>% 
  left_join(attrib_hosps_for_summary_nogeo, by = "zcta") %>% 
  filter(is.na(attrib_hosps)==F) %>% 
  dplyr::select(zcta, contains("attrib_hosps")) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "attrib_hosps",
    lwd=0.05,
    fill.legend= tm_legend("Heat-attributable hospitalizations per heat-wave day"),
    fill.scale=tm_scale_intervals(
      style="pretty",
      values="matplotlib.viridis")
  )

#tmap_attrib_hosps

#tmap is failing for some reason. try mapview instead
mv_tmap_attrib_hosps= zcta_ca_geo_simplified %>% 
  dplyr::select(-contains("pop")) %>% 
  dplyr::select(-contains("zcta_in_cali")) %>% 
  left_join(attrib_hosps_for_summary_nogeo, by = "zcta") %>% 
  filter(is.na(attrib_hosps)==F) %>% 
  dplyr::select(zcta, contains("attrib_hosps")) %>% 
  mapview(zcol="attrib_hosps",
          layer.name="Heat-attributable hospitalizations",
          lwd=0.1)
mv_tmap_attrib_hosps

```

## Attributable number of hospitalizations per heat-wave day (baseline) per population
```{r}
#| echo: false
#| warning: false
#| message: false
tmap_attrib_hosps_per_pop= zcta_ca_geo_simplified %>% 
  dplyr::select(-contains("pop")) %>% 
  dplyr::select(-contains("zcta_in_cali")) %>% 
  left_join(attrib_hosps_for_summary_nogeo, by = "zcta") %>% 
  filter(is.na(attrib_hosps)==F) %>% 
  dplyr::select(zcta, contains("attrib_hosps")) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "attrib_hosps_per_pop",
    lwd=0.05,
    fill.legend= tm_legend("Heat-attributable hospitalizations per heat-wave day per population"),
    fill.scale=tm_scale_intervals(
      style="pretty",
      values="matplotlib.viridis")
  )
  

#tmap_attrib_hosps_per_pop

#tmap is failing for some reason. try mapview instead
mv_tmap_attrib_hosps_per_pop= zcta_ca_geo_simplified %>% 
  dplyr::select(-contains("pop")) %>% 
  dplyr::select(-contains("zcta_in_cali")) %>% 
  left_join(attrib_hosps_for_summary_nogeo, by = "zcta") %>% 
  filter(is.na(attrib_hosps_per_pop)==F) %>% 
  #remove infinity
  filter(attrib_hosps_per_pop<100) %>% 
  dplyr::select(zcta, contains("attrib_hosps_per_pop")) %>% 
  mapview(zcol="attrib_hosps_per_pop",
          layer.name="Heat-attributable hospitalizations per pop.",
          lwd=0.1)
mv_tmap_attrib_hosps_per_pop

```

## Tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false
tmap_mode("view")  
#Use tmap instead of qtm
tmap_tree= zcta_ca_geo_simplified %>% 
  left_join(covars, by = "zcta") %>% 
  #remove everything without a geometry
  filter(is.na(TreeCanopy)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "TreeCanopy",
    lwd=0.5,
    #Change legend. This took some doing to get right.
    #Use X.legend where X corresponds to the layer being
    #visualized, fill here
    #See https://r-tmap.github.io/tmap/reference/qtm.html
    fill.legend= tm_legend("Tree canopy share"),
    
    #This is how to change the colors. select a viridis scale
    #https://r-tmap.github.io/tmap/articles/adv_shiny
    #within matplotlib
    fill.scale=tm_scale_intervals(
            #keep number of categories fewer. see
      #https://r-tmap.github.io/tmap/articles/basics_scales

      style="pretty",
      values="brewer.yl_gn")
) 

tmap_tree
```

## Impervious surface
```{r}
#| echo: false
#| warning: false
#| message: false
tmap_impsurf= zcta_ca_geo_simplified %>% 
  left_join(covars, by = "zcta") %>% 
  filter(is.na(ImpervSurf)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "ImpervSurf",
    lwd=0.5,
    fill.legend= tm_legend("Impervious surfaces"),
    fill.scale=tm_scale_intervals(
      style="pretty",
      values="brewer.reds")
) 

tmap_impsurf
```
## SVI
```{r}
#| echo: false
#| warning: false
#| message: false
tmap_svi= zcta_ca_geo_simplified %>% 
  left_join(covars, by = "zcta") %>% 
  filter(is.na(SVI)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "SVI",
    lwd=0.5,
    fill.legend= tm_legend("SVI"),
    fill.scale=tm_scale_intervals(
      style="pretty",
      values="brewer.purples")
) 

tmap_svi

```

# IRR

## Impervious surfaces

### Population-based

```{r}
#| echo: false
#| warning: false
#| message: false
#impervious surfaces
irr_imp_pb =  irr_long %>% 
  filter(scenario_intervention=="Imp") %>% 
  filter(scenario_type_7=="Pop. based") %>% 
  dplyr::select("zcta", starts_with("scenario"), "measure", starts_with("value")) 

tmap_mode("view") #run again just in case it was changed above.


#see below for defining palette
#https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/qtm
tmap_irr_imp_pb= zcta_ca_geo_simplified %>% 
  left_join(irr_imp_pb, by = "zcta") %>% 
  #remove everything without a geometry
  filter(is.na(value)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
    tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "value_cat",
    lwd=.001,
    fill.legend= tm_legend("IRR"),
    #This is technically categorical, as I
    #already divided it up
    fill.scale=tm_scale_categorical(
#      values="brewer.rd_bu" #also tried brewer.rd_bu
      #Note all you have to do flip the palette is put a minus
      #sign in front of it
      values="-brewer.rd_yl_bu"
    )
 )

tmap_irr_imp_pb
```

### Proportionate Universalism 1

```{r}
#| echo: false
#| warning: false
#| message: false
#impervious surfaces
irr_imp_pu1 =  irr_long %>% 
  filter(scenario_intervention=="Imp") %>% 
  filter(scenario_type_7=="Prop. Univ. 1") %>% 
  dplyr::select("zcta", starts_with("scenario"), "measure", starts_with("value")) 

tmap_irr_imp_pu1= zcta_ca_geo_simplified %>% 
  left_join(irr_imp_pu1, by = "zcta") %>% 
  #remove everything without a geometry
  filter(is.na(value)==F) %>% 
    tm_shape()+ #begin with tmap shape. pipe in the above.
    tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "value_cat",
    lwd=.001,
    fill.legend= tm_legend("IRR"),
    #This is technically categorical, as I
    #already divided it up
    fill.scale=tm_scale_categorical(
#      values="brewer.rd_bu" #also tried brewer.rd_bu
      #Note all you have to do flip the palette is put a minus
      #sign in front of it
      values="-brewer.rd_yl_bu"
    )
 )
  
tmap_irr_imp_pu1
```

### Proportionate Universalism 2

```{r}
#| echo: false
#| warning: false
#| message: false
#impervious surfaces
#table(irr_long$scenario_type_7)
irr_imp_pu2 =  irr_long %>% 
  filter(scenario_intervention=="Imp") %>% 
  filter(scenario_type_7=="Prop. Univ. 2") %>% 
  dplyr::select("zcta", starts_with("scenario"), "measure", starts_with("value")) 

tmap_irr_imp_pu2= zcta_ca_geo_simplified %>% 
  left_join(irr_imp_pu2, by = "zcta") %>% 
  filter(is.na(value)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "value_cat",
    lwd=.001,
    fill.legend= tm_legend("IRR"),
    fill.scale=tm_scale_categorical(
      values="-brewer.rd_yl_bu"
    )
 )

tmap_irr_imp_pu2
```

### Proportionate Universalism 3

```{r}
#| echo: false
#| warning: false
#| message: false
#impervious surfaces
#table(irr_long$scenario_type_7)
irr_imp_pu3 =  irr_long %>% 
  filter(scenario_intervention=="Imp") %>% 
  filter(scenario_type_7=="Prop. Univ. 3") %>% 
  dplyr::select("zcta", starts_with("scenario"), "measure", starts_with("value")) 

tmap_irr_imp_pu3= zcta_ca_geo_simplified %>% 
  left_join(irr_imp_pu3, by = "zcta") %>% 
  filter(is.na(value)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "value_cat",
    lwd=.001,
    fill.legend= tm_legend("IRR"),
    fill.scale=tm_scale_categorical(
      values="-brewer.rd_yl_bu"
    )
 )

tmap_irr_imp_pu3
```

## Tree canopy

# IRD

## Impervious surfaces

### Population-based

```{r}
#| echo: false
#| warning: false
#| message: false
#impervious surfaces
ird_imp_pb =  ird_long %>% 
  filter(scenario_intervention=="Imp") %>% 
  filter(scenario_type_7=="Pop. based") %>% 
  dplyr::select("zcta", starts_with("scenario"), "measure", starts_with("value")) 

tmap_mode("view") #run again just in case it was changed above.

#see below for defining palette
#https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/qtm
tmap_ird_imp_pb= zcta_ca_geo_simplified %>% 
  left_join(ird_imp_pb, by = "zcta") %>% 
  filter(is.na(value)==F) %>% 
  tm_shape()+ #begin with tmap shape. pipe in the above.
  tm_basemap("CartoDB.Positron")+#define the basemap
  tm_polygons(
    fill = "value_cat",
    lwd=.001,
    fill.legend= tm_legend("IRD"),
    fill.scale=tm_scale_categorical(
      values="matplotlib.viridis"
    )
 )

tmap_ird_imp_pb
```