---
title: "Static maps"
author: "Michael D Garber"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
---

Create faceted static maps using ggplot2

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
#March 13, 2025: updating in the more general climate-health project folder
library(readr)
library(readxl)
library(here)
library(tidyverse)
library(viridis)
library(sf)
library(maptiles)#to get basemap
library(tidyterra) #to plot the basemap
#load measures
source(
  here("scripts","scripts-heat-hosp-green-hia","read-attrib-hosps-ratios-diffs.R"))
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")
load("california_boundaries_sf.RData")

#get tiles for basemap
california_bmap_cartodbpositron=get_tiles(
  california_boundaries_sf,  provider = "CartoDB.Positron", crop = TRUE)
```



# IRR

## Population-based scenarios: Impervious surfaces vs. Tree canopy
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
ggplot_facet_pb_imp_v_tree=function(df,measure_value, direction_value){
  df %>% 
  ggplot()+
  #March 13, 2025: add a basemap
  geom_spatraster_rgb(data = california_bmap_cartodbpositron) +
  geom_sf(
    aes(fill=value_cat),
    linewidth=.05)+
  #change palette to brewer diverging
  # scale_fill_brewer(
  #   type="div",
  #   palette = "RdYlBu",
  #   direction=direction_value
  # )+
  scale_fill_viridis(
    direction=direction_value, #flip direction for ratios
    discrete=TRUE,
    name=measure_value, #legend title
    na.value="grey")+
  # theme_classic()+
  theme_minimal()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )+
  facet_grid(
    cols=vars(scenario_intervention)
  )
}

gg_irr_pb=zcta_ca_geo_simplified %>% 
  left_join(irr_long, by = "zcta") %>% 
  filter(scenario_type_3 =="Population-based") %>% 
  ggplot_facet_pb_imp_v_tree(measure_value="IRR",direction_value=-1)

gg_irr_pb
```


## Proportionate Universalism and Targeted Scenarios
### Impervious Surfaces

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
ggplot_facet_each_int_type=function(df, measure_value, direction_value){
  df %>% 
  filter(scenario_type_3 !="Population-based") %>% #excluding these
  ggplot()+
  geom_spatraster_rgb(data = california_bmap_cartodbpositron) +
  geom_sf(
    aes(fill=value_cat),
    linewidth=.05)+
  #change palette to brewer diverging
    #makes clearer when ratio is above 1
  # scale_fill_brewer(
  #   type="div",
  #   palette = "RdYlBu",
  #   direction=direction_value
  # )+
  scale_fill_viridis(
    direction=direction_value, #flip direction for ratios
    discrete=TRUE,
    name=measure_value, #legend title
    na.value="grey")+
#  theme_classic()+
  theme_minimal()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
    )+
  facet_grid(
    cols=vars(scenario_sub_type),
    rows=vars(scenario_type_3)
  )
  
}

zcta_ca_geo_simplified %>% 
  left_join(irr_long, by = "zcta") %>% 
  filter(scenario_intervention=="Imp") %>% 
  ggplot_facet_each_int_type(measure_value="IRR",direction_value=-1)


```

### Tree canopy

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(irr_long, by = "zcta") %>% 
  filter(scenario_intervention=="Tree") %>% 
  ggplot_facet_each_int_type(measure_value="IRR",direction_value=-1)
```

# IRD

## Population-based scenarios: Impervious surfaces vs. Tree canopy

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(ird_long, by = "zcta") %>% 
  filter(scenario_type_3 =="Population-based") %>% 
  ggplot_facet_pb_imp_v_tree(measure_value="IRD",direction_value=1)
```

### Impervious surfaces

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(ird_long, by = "zcta") %>% 
  filter(scenario_intervention=="Imp") %>% 
  ggplot_facet_each_int_type(measure_value="IRD",direction_value=1)
```

### Tree canopy
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(ird_long, by = "zcta") %>% 
  filter(scenario_intervention=="Tree") %>% 
  ggplot_facet_each_int_type(measure_value="IRD",direction_value=1)
```


# Difference, number of cases

## Population-based scenarios: Impervious surfaces vs. Tree canopy

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(n_cases_diff_long, by = "zcta") %>% 
  filter(scenario_type_3 =="Population-based") %>% 
  ggplot_facet_pb_imp_v_tree(measure_value="n_cases_diff",direction_value=1)
```

### Impervious surfaces

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(n_cases_diff_long, by = "zcta") %>% 
  filter(scenario_intervention=="Imp") %>% 
  ggplot_facet_each_int_type(measure_value="n_cases_diff",direction_value=1)
```

### Tree canopy
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
zcta_ca_geo_simplified %>% 
  left_join(n_cases_diff_long, by = "zcta") %>% 
  filter(scenario_intervention=="Tree") %>% 
  ggplot_facet_each_int_type(measure_value="n_cases_diff",direction_value=1)
```

