---
title: "results"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
---

```{r}
#| echo: false
#| warning: false
library(here)
library(mapview) #make sure mapview is github version
library(tidyverse)
library(viridis)
library(viridisLite)
library(leaflet)
library(leaflet.extras)
library(tmap)
library(knitr)
library(RColorBrewer)
source(here("scripts","read-wrangle-eff-estimates.R"))#load measures
```

# Introduction
Effects of wildfire smoke on daily respiratory acute-care utilization has been estimated at the zip-code level in California (Do et al).

Rate differences per 100,000 are estimated at the zip code level.

Effect modification of this effect by community characteristics has been estimated.

The increase in risk difference per IQR increase in air conditioning prevalence is -0.239302618 (95% CI, -0.41143431, -0.0671709235), the corresponding IQR for AC prevalence is 6.000915e-01.

Over-arching question: 
How would hypothetically changing the distribution of the effect modifiers affect the total burden and spatial distribution of respiratory acute-care utilization?

## Baseline risk difference (zcta-level)
### Histogram and summary statistics
RD is per 100,000
```{r}
#| echo: false
#| warning: false

setwd(here("data-processed"))
load("wf_ac_zcta_wrangle.RData")
summary(wf_ac_zcta_wrangle$rd_baseline_pt)
wf_ac_zcta_wrangle %>% 
  ggplot(aes(x=rd_baseline_pt))+
  geom_histogram()+
  theme_bw()
```


### Map
```{r}
#| echo: false
#| warning: false
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")

ac_zcta_among_wf=wf_ac_zcta %>% 
  dplyr::select(zcta,ac_prop)
tmap_mode("view") 
qtm_rd_baseline= zcta_ca_geo_simplified %>% 
  left_join(wf_ac_zcta_wrangle, by = "zcta") %>% 
  filter(is.na(rd_baseline_pt)==F) %>% 
  qtm("rd_baseline_pt")
qtm_rd_baseline
```

## Air conditioning (zcta-level)
Distribution of air conditioning at the zip-code level

### Histogram and summary statistics
```{r}
#| echo: false
#| warning: false

summary(wf_ac_zcta$ac_prop)
wf_ac_zcta %>% 
  ggplot(aes(x=ac_prop))+
  geom_histogram()+
  theme_bw()
```

### Map
```{r}
#| echo: false
#| warning: false
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")

ac_zcta_among_wf=wf_ac_zcta %>% 
  dplyr::select(zcta,ac_prop)
tmap_mode("view") 
qtm_ac= zcta_ca_geo_simplified %>% 
  left_join(ac_zcta_among_wf, by = "zcta") %>% 
  filter(is.na(ac_prop)==F) %>% 
  qtm("ac_prop",fill.palette="YlOrRd")
qtm_ac
```

# Methods
## Scenario definitions:
Hypothetically change the distribution of AC in three ways to assess how it would change the risk difference at the ZCTA level.

```{r}
#| echo: false
#| warning: false
ac_iqr_among_wf_eff %>% 
  kable()
```


Scenario 1: Among those zip codes with an AC prevalence below the median, raise their AC prevalence to the median (0.6250).

Scenario 2: Among those zip codes with an AC prevalence below the 75th percentile, raise their AC prevalence to the 75th percentile (0.8710).

Scenario 3: Same, but 100th percentile (1).

## HIA method
For each zip code, calculate the change in the AC proportion from the status quo to the target level. For example, if a zip code has a 50% AC prevalence, the difference from baseline to target in Scenario 1 is 12.5% (65.5%-50%).
Then, express that difference in terms of the number of IQRs that it represents. The IQR of AC is 0.61 (above).
That zip code would raise its AC prevalence by
```{r}
0.125/.61
```

Then, use that value to calculate the new risk difference under that scenario, following this equation:

rd_target_pt=rd_baseline_pt+rd_per_ac_iqr_pt*ac_prop_change_per_iqr

where

* rd_target_pt = the zip code's new risk difference under the scenario

* rd_per_ac_iqr_pt = the increase in the risk difference per change in IQR of AC

* ac_prop_change_per_iqr = the number of IQRs changed in that zip code in that scenario

This assumes that the increase is linear.

## Uncertainty 
In each of 1,000 replicatess, re-sample **rd_baseline_pt** and **rd_per_ac_iqr_pt** from a normal distribution using the reported standard deviation and 95% CI (assuming confidence limit is estimate +/- 1.96*SD).

Then, in each replicate, propagate this uncertainty by re-calculating **rd_target_pt** using the re-sampled input values.

The resulting uncertainty interval is 2.5th and 97.5th percentiles.


# Results
## Overall summary

There are 1,396 total zip codes

**n_zcta_intervene** is the number of zip codes whose AC values would be changed under the scenario.

Risk differences are expressed per 100,000.

Means and sums are for those zctas with intervention
```{r}
#| echo: false
#| warning: false

setwd(here("data-processed"))
load("all_scenarios_summary_pt.RData")
load("all_scenarios_summary_boot.RData")
all_scenarios_summary_pt %>% 
  dplyr::select(
    target_percentile,
    n_zcta_intervene,
    rd_baseline_sum,
    rd_baseline_mean_unwt,
    rd_target_sum,
    rd_target_mean_unwt,
    rd_target_v_baseline_diff_in_sum
  ) %>% 
  left_join(all_scenarios_summary_boot,by="target_percentile") %>% 
  dplyr::select(
    contains("target_percentile"),
        n_zcta_intervene,
    contains("rd_baseline_sum"),
    contains("rd_baseline_mean"),
    contains("rd_target_sum"), 
    contains("rd_target_mean"),
    contains("rd_target_v_baseline_diff_in_sum")
  ) %>% 
  knitr::kable(digits=3)
```


## Maps of results for each scenario
Values mapped are the risk difference per 100k for the corresponding scenario among those zip-codes that were intervened upon in that scenario.

### Target percentile: 0.5
```{r}
#| echo: false
#| warning: false
setwd(here("data-processed"))
load("all_scenarios_summary_by_zcta_pt.RData")

tmap_mode("view") 

all_scenarios_summary_by_zcta_pt_scenario_1=all_scenarios_summary_by_zcta_pt %>% 
  filter(target_percentile==0.5)


qtm_rd_sc_1= zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt_scenario_1, by = "zcta") %>% 
  filter(is.na(rd_baseline_mean_unwt)==F) %>% 
  qtm("rd_baseline_mean_unwt")
qtm_rd_sc_1
```

### Target percentile: 0.75
```{r}
#| echo: false
#| warning: false
setwd(here("data-processed"))
load("all_scenarios_summary_by_zcta_pt.RData")

tmap_mode("view") 

all_scenarios_summary_by_zcta_pt_scenario_2=all_scenarios_summary_by_zcta_pt %>% 
  filter(target_percentile==0.75)


qtm_rd_sc_2= zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt_scenario_2, by = "zcta") %>% 
  filter(is.na(rd_baseline_mean_unwt)==F) %>% 
  qtm("rd_baseline_mean_unwt")
qtm_rd_sc_2
```


### Target percentile: 1
```{r}
#| echo: false
#| warning: false
setwd(here("data-processed"))
load("all_scenarios_summary_by_zcta_pt.RData")

tmap_mode("view") 

all_scenarios_summary_by_zcta_pt_scenario_3=all_scenarios_summary_by_zcta_pt %>% 
  filter(target_percentile==1)


qtm_rd_sc_3= zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt_scenario_3, by = "zcta") %>% 
  filter(is.na(rd_baseline_mean_unwt)==F) %>% 
  qtm("rd_baseline_mean_unwt")
qtm_rd_sc_3
```

