---
title: "Results: Modification of effect of wildfire on acute-care utilization by air-conditioning at the zip-code level"
author: "Michael D Garber"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
---

# Technical notes
Using rmarkdown because quarto and tmap are not working well together at the moment.

URL of this page: https://michaeldgarber.github.io/climate-health/results-markdown.html

Some related in-progress Shiny apps, which may not be fully functional:

* An app with interactive (leaflet) maps displaying how the difference in risk differences between target and baseline scenarios varies over scenarios. Currently this works with mapview, but not tmap. They are similar mapping packages. For this purpose, I prefer tmap as it provides clearer binning of values, so will keep working on why interactive tmaps are not working on the server. https://michaeldgarber.shinyapps.io/tmap-interactive/

* Same as above but using static (non-interactive) maps. Note that I am currently working on how to get this shiny app to display on the server with a basemap. I can currently get this to work with a basemap locally on my computer, but it fails upon pushing to the server. https://michaeldgarber.shinyapps.io/tmap-static-lite/

* An app comparing the difference in risk differences between two selected scenarios (e.g., 80th percentile with 30th percentile): https://michaeldgarber.shinyapps.io/compare-scenarios/




```{r}
#| echo: false
#| warning: false
#| message: false
library(here)
library(mapview) #make sure mapview is github version
library(tmap)
library(leafgl)
library(tidyverse)
library(viridis)
library(viridisLite)
library(leaflet)
library(leaflet.extras)
library(tmap)
library(knitr)
library(RColorBrewer)
source(here("scripts","read-wrangle-eff-estimates.R"))#load measures
```

# Introduction and data exploration
Effects of wildfire smoke on daily respiratory acute-care utilization has been estimated at the zip-code level in California (Do et al).

Rate differences per 100,000 are estimated at the zip code level.

Effect modification of this effect by community characteristics has been estimated.

The increase in risk difference per IQR increase in air conditioning prevalence is -0.239302618 (95% CI, -0.41143431, -0.0671709235), the corresponding IQR for AC prevalence is 6.000915e-01.

Over-arching question: 
How would hypothetically changing the distribution of the effect modifiers affect the total burden and spatial distribution of respiratory acute-care utilization?

## Baseline risk difference (zcta-level)
### Histogram and summary statistics
RD is per 100,000
```{r}
#| echo: false
#| warning: false
#| message: false

setwd(here("data-processed"))
load("wf_ac_zcta_wrangle.RData")
summary(wf_ac_zcta_wrangle$rd_baseline_pt)
wf_ac_zcta_wrangle %>% 
  ggplot(aes(x=rd_baseline_pt))+
  geom_histogram()+
  theme_bw()+
  xlab("Risk difference per 100,000 (baseline)")
```


### Map of baseline risk difference
```{r}
#| echo: false
#| warning: false
#| message: false

setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")

ac_zcta_among_wf=wf_ac_zcta %>% 
  dplyr::select(zcta,ac_prop)
tmap_mode("view") 
qtm_rd_baseline= zcta_ca_geo_simplified %>% 
  left_join(wf_ac_zcta_wrangle, by = "zcta") %>% 
  dplyr::select(
    contains("zcta"),
    contains("pop"),
    contains("rd_baseline"),
    contains("ac_prop")
    ) %>% 
  filter(is.na(rd_baseline_pt)==F) %>% 
  qtm(
    #choose just 1 basemap to decrease file size
    basemaps = "CartoDB.Positron",
        lwd=.5,
    fill = "rd_baseline_pt",
    title = "Risk difference (baseline)",
      #Change legend. This took some doing to get right.
      #Use X.legend where X corresponds to the layer being
      #visualized, fill here
      #See https://r-tmap.github.io/tmap/reference/qtm.html
      fill.legend= tm_legend("Risk Difference (per 100k)")
)+
  tm_view()

qtm_rd_baseline
```


## Uncertainty and precision in baseline RD
Explore uncertainty in the baseline risk difference estimates across ZCTAS.
Consider how uncertainty may vary across zip codes by such things as zip code population or urban-rural status.
Consider limiting analyses to only those ZCTAS whose baseline risk differences are sufficiently precise.

### SD in baseline Risk difference by zip-code rural-urban commmuting code
Definitions of rural-urban commuting codes:
1. Metropolitan area core: primary flow within an urbanized area (UA)
2. Metropolitan area high commuting: primary flow 30% or more to a UA
3. Metropolitan area low commuting: primary flow 10% to 30% to a UA
4. Micropolitan area core: primary flow within an Urban Cluster of 10,000 to 49,999 (large UC)
5. Micropolitan high commuting: primary flow 30% or more to a large UC
6. Micropolitan low commuting: primary flow 10% to 30% to a large UC
7. Small town core: primary flow within an Urban Cluster of 2,500 to 9,999 (small UC)
8. Small town high commuting: primary flow 30% or more to a small UC
9. Small town low commuting: primary flow 10% to 30% to a small UC
10. Rural areas: primary flow to a tract outside a UA or UC



```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("zcta_ruca2010.RData")
load("wf_ac_zcta_wrangle.RData")
wf_ac_zcta_ruca_wrangle=wf_ac_zcta_wrangle %>% 
  left_join(zcta_ruca2010,by="zcta")
#alright...so, yes, the estimates are more precise
#for the metro areas
wf_ac_zcta_ruca_wrangle %>% 
  ggplot(aes(x=ruca_cat,y=rd_baseline_sd))+
  geom_boxplot()+
  labs(
    y="SD in Baseline risk difference, ZCTA",
    x="Rural-Urban Commuting Code (defined above)"
  )+
  theme_bw()
```

Interpretation: the estimates are more precise in more urban zip codes.
That follows as they'd have a higher population.

### SD in baseline risk difference plotted against zip-code population
How does population itself relate to the standard deviation of the baseline risk-difference estimate?

Uncertainty in the estimate decreases with increasing ZCTA population.
```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("lookup_pop_zcta.RData")
wf_ac_zcta_ruca_wrangle %>% 
  left_join(lookup_pop_zcta,by="zcta") %>% 
  ggplot(aes(x=pop,y=rd_baseline_sd))+
  geom_point()+
  geom_smooth()+
  labs(
    y="SD in Baseline risk difference, ZCTA",
    x="Population of ZCTA (2020)"
  )+
  theme_bw()
```


### Map of standard error of risk difference
To see spatial distribution of precision in effect estimates.
As above, more precise values are in urban centers such as SF Bay, LA, and SD.
```{r}
#| echo: false
#| warning: false
#| message: false

setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")

ac_zcta_among_wf=wf_ac_zcta %>% 
  dplyr::select(zcta,ac_prop)
tmap_mode("view") 
qtm_rd_sd_baseline= zcta_ca_geo_simplified %>% 
  left_join(wf_ac_zcta_wrangle, by = "zcta") %>% 
  dplyr::select(
    contains("zcta"),
    contains("pop"),
    contains("rd_baseline"),
    contains("ac_prop")
    ) %>% 
  filter(is.na(rd_baseline_pt)==F) %>% 
  qtm(
    #choose just 1 basemap to decrease file size
    basemaps = "CartoDB.Positron",
    lwd=.25,
    
    fill = "rd_baseline_sd",
    title = "Standard error in risk difference (baseline)",
      #See https://r-tmap.github.io/tmap/reference/qtm.html
      fill.legend= tm_legend("Standard error in risk difference (per 100k)")
)+
  tm_view()

qtm_rd_sd_baseline
```

## Air conditioning (zcta-level)
Distribution of air conditioning at the zip-code level

### Histogram and summary statistics
```{r}
#| echo: false
#| warning: false
#| message: false
summary(wf_ac_zcta$ac_prop)
wf_ac_zcta %>% 
  ggplot(aes(x=ac_prop))+
  geom_histogram()+
  theme_bw()+
  xlab("Proportion of residents with AC")+
  ylab("Number of ZCTAs")
```

### Map of air conditioning
```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")

ac_zcta_among_wf=wf_ac_zcta %>% 
  dplyr::select(zcta,ac_prop)
tmap_mode("view") 
qtm_ac= zcta_ca_geo_simplified %>% 
  left_join(ac_zcta_among_wf, by = "zcta") %>% 
    qtm(
      fill = "ac_prop",
      title = "Proportion with A/C",
      fill.legend= tm_legend("Proportion with A/C"),
      
      #choose just 1 basemap to decrease file size
      basemaps = "CartoDB.Positron"
)

qtm_ac %>% 
  leaflet.extras::addFullscreenControl()
```



# Methods


```{r}
#| echo: false
#| warning: false
#| message: false
ac_iqr_among_wf_eff %>% 
  kable()
```



## HIA method
For each zip code, calculate the change in the AC proportion from the status quo to the target percentile level. For example, if a zip code has a 50% AC prevalence, the difference from baseline to target in Scenario 1 is 12.5% (62.5%-50%).

Then, express that difference in terms of the number of IQRs that it represents. The IQR of AC is 0.61 (above).
That zip code would raise its AC prevalence by
```{r}
0.125/.61
```

Then, use that value to calculate the new risk difference under that scenario, following this equation:

rd_target_pt=rd_baseline_pt+rd_per_ac_iqr_pt*ac_prop_change_per_iqr

where

* rd_target_pt = the zip code's new risk difference under the scenario

* rd_per_ac_iqr_pt = the increase in the risk difference per change in IQR of AC

* ac_prop_change_per_iqr = the number of IQRs changed in that zip code in that scenario

This assumes that the increase is linear.
It also assumes no confounding.

## Scenario definitions:
Considering 20 scenarios for 20 possible target percentile values (0th, 5th,....95th, 100th)

## Uncertainty 
In each of 1,000 replicatess, re-sample **rd_baseline_pt** and **rd_per_ac_iqr_pt** from a normal distribution using the reported standard deviation and 95% CI. Assume the confidence limit is the estimate +/- 1.96*SD.

Then, in each replicate, propagate this uncertainty throughout the calculations by re-calculating **rd_target_pt** using the re-sampled input values.

The resulting uncertainty interval is the 2.5th and 97.5th percentiles.


# Results
## Overall summary

### Table
There are **1,396** total zip codes.

The right-most columns present the difference in risk differences between the target scenario and the baseline scenario (i.e., a difference in risk differences).


```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("all_scenarios_summary_pt.RData")
load("all_scenarios_summary_boot.RData")
#Comment: the confidence intervals seem backwards.
all_scenarios_summary_pt %>% 
  dplyr::select(
    target_percentile,
    n_zcta_intervene,
    rd_baseline_sum,
    rd_target_sum,
    rd_diff_sum
  ) %>% 
  left_join(all_scenarios_summary_boot,by="target_percentile") %>% 
  dplyr::select(
    contains("target_percentile"),
        n_zcta_intervene,
    contains("rd_baseline_sum"),
    contains("rd_target_sum"), 
    contains("rd_diff_sum")
  ) %>% 
  knitr::kable(
    col.names = c(
      "Target percentile","N, ZCTAS intervened",
      "RD per 100k, baseline (sum over ZCTAs)",
      "RD per 100k, baseline (sum over ZCTAs) (LL)",
      "RD per 100k, baseline (sum over ZCTAs) (UL)",
      "RD per 100k, target (sum over ZCTAs)",
      "RD per 100k, target (sum over ZCTAs) (LL)",
      "RD per 100k, target (sum over ZCTAs) (UL)",
      "RD per 100k, difference (sum over ZCTAs)",
      "RD per 100k, difference (sum over ZCTAs) (LL)",
      "RD per 100k, difference (sum over ZCTAs) (UL)"
      ),
    digits=2)
    
```

### Plot of difference in risk difference by scenario 
A plot of the "difference in risk differences" column in the above table.
The vertical error bars denote the 95% uncertainty interval.
```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("all_scenarios_summary_pt.RData")
load("all_scenarios_summary_boot.RData") 
#comment: add uncertainty intervals to this

all_scenarios_summary_pt  %>%
  left_join(all_scenarios_summary_boot,by="target_percentile") %>% 
  dplyr::select(
    contains("target_percentile"),
        n_zcta_intervene,
    contains("rd_baseline_sum"),
    contains("rd_target_sum"), 
    contains("rd_diff_sum")
  ) %>% 
  ggplot(aes(x=target_percentile,y=rd_diff_sum))+
  geom_point()+
  geom_pointrange(
        aes(
        ymin=rd_diff_sum_ll,
        ymax=rd_diff_sum_ul),

    fatten=1,
    )+
  labs(
    x="Target percentile",
    y="Difference in risk differences\n(per 100,000)\n(target scenario-baseline)"
  )+
  theme_bw()
```


## How distribution changes by scenario

### Distribution of target risk difference by scenario (histogram panel)
Note this only includes every other scenario to facilitate clearer visualization.

Note x-axis is restricted to below 7 and above -7.
How many does that exclude?
```{r}
#how many are out of range if I censor x-axis at -7 and 7?
setwd(here("data-processed"))
load("wf_ac_scenarios_all_pt.RData")
wf_ac_scenarios_all_pt %>% 
  filter(target_percentile_decile==1) %>% 
  mutate(rd_target_pt_cat=case_when(
    rd_target_pt>7~"7+",
    rd_target_pt<=7 & rd_target_pt>-7~"-7 to 7",
    rd_target_pt<=-7 ~"< -7"
  )) %>% 
  group_by(rd_target_pt_cat) %>% 
  summarise(
    n=n()) %>% 
  ungroup() %>% 
  mutate(prop=n/sum(n))
```

Restricted to target RD above -7 and below 7 to more clearly see the meaningful range.
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 10 #make taller so easier to see

setwd(here("data-processed"))
load("wf_ac_scenarios_all_pt.RData")

  
wf_ac_scenarios_all_pt %>% 
  #to slim down the number of charts
  filter(target_percentile_decile==1) %>% 
  ggplot(aes(x=rd_target_pt, fill = zcta_intervene_char_yn))+
  geom_histogram(binwidth = .1)+
  facet_grid(rows=vars(target_percentile))+
  theme_bw(base_size = 16)+
  #add limits on the x-axis so the bulk of the histogram
  #in the middle is more visible
  scale_x_continuous(
    limits=c(-7,7)
  )+
  labs(
    y="Number of ZCTAs",
    x="Target risk difference per 100,000",
    fill="ZCTA intervened upon")
```


### Distribution of difference in risk differences (histogram panel)
Comparing the target scenario with baseline.
Note this only includes every other scenario to facilitate clearer visualization
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 10 #make taller so easier to see
wf_ac_scenarios_all_pt %>% 
  filter(target_percentile_decile==1) %>% 
  ggplot(aes(x=rd_diff_pt, fill = zcta_intervene_char_yn))+
  #make bins shorter to make the visualization clearer
  #.2 is too coarse;
  geom_histogram(binwidth=.05)+
  facet_grid(rows=vars(target_percentile))+
  theme_bw(base_size = 16)+
    labs(
    y="Number of ZCTAs",
    #using \n for line break
    x="Difference in risk differences\nper 100,000\n(Target-Baseline)",
    fill="ZCTA intervened upon")
```

### Distribution of difference in risk differences (boxplot panel)
Diamond points are the mean over ZCTAs. Horizontal line is the median. Top and bottom of box are 25th and 75th percentiles.
```{r}
#| echo: false
#| warning: false
#| message: false
wf_ac_scenarios_all_pt %>%
  mutate(target_percentile_char=as.character(target_percentile)) %>% 
#  filter(zcta_intervene==1) %>% 
  filter(target_percentile_decile==1) %>% 
  ggplot(aes(x= target_percentile_char, y = rd_diff_pt))+
  geom_boxplot(
    outlier.size =  .01,
    outlier.colour = "gray",
    varwidth=TRUE,
  )+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  theme_bw(base_size=14)+
  labs(
    x="Target percentile",
    y="Difference in risk differences\nper 100,000\n(Target-Baseline)"
  )
```

## Facet map: change in spatial distribution
Show how the distribution changes spatially over the scenarios
```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 10 #make taller so easier to see
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")
load("lookup_target_percentile_decile.RData")
load("all_scenarios_summary_by_zcta_pt.RData")
tmap_mode("plot") 
#This creates a faceted map. Also see this chapter by Lovelace
#https://r.geocompx.org/adv-map#faceted-maps
#Note I'm using the conventional tmap syntax here rather than qtm,
#which stands for quick thematic map.
zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt,by="zcta") %>% 
  left_join(lookup_target_percentile_decile,
            by="target_percentile") %>% 
  filter(target_percentile_decile==1) %>% 
  tm_shape()+
  #for specifying the color scale
  #https://r-tmap.github.io/tmap/articles/basics_scales
  tm_polygons(
    fill="rd_diff_sum",
    #c is Plasma
    palette=viridis(n=4, option="C"),#consistent with above
    title="Difference in risk differences\nper 100,000\n(Target-Baseline)",
    lwd=0.1)+
  tm_facets(by="target_percentile" 
            #nrow=10, 
            #scale.factor = 1
            )+
  #add a basemap at the end of the code like this:
  #https://r-tmap.github.io/tmap/reference/tm_basemap.html
  tm_basemap("CartoDB.PositronNoLabels") 


```


## Maps (interactive) of results for selected scenarios
Values mapped are the risk difference per 100k for the corresponding scenario among those zip-codes that were intervened upon in that scenario.

A total of 20 scenarios are available. Choose two, 0.75 and 1.

Note to self: keep color scales constant between maps to facilitate comparison.

### Target percentile: 0.75
```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("all_scenarios_summary_by_zcta_pt.RData")


all_scenarios_summary_by_zcta_pt_scenario_75=all_scenarios_summary_by_zcta_pt %>% 
  filter(target_percentile==0.75) %>% 
  #and only pick those to print
  dplyr::select(zcta,
                n_zcta_intervene,
                rd_baseline_sum,
                rd_target_sum,
                rd_diff_sum
                )

tmap_mode("view") 
tm_75_obj= zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt_scenario_75, by = "zcta") 

qtm_rd_sc_75=tm_basemap("CartoDB.Positron")+
  tm_shape(tm_75_obj)+
  tm_polygons(
    fill = "rd_diff_sum",
    lwd=0.5,
    #Change legend. This took some doing to get right.
    #Use X.legend where X corresponds to the layer being
    #visualized, fill here
    #See https://r-tmap.github.io/tmap/reference/qtm.html
    fill.legend= tm_legend("Difference in risk differences (per 100k)"),
    
    #This is how to change the colors. select a viridis scale
    #https://r-tmap.github.io/tmap/articles/adv_shiny
    #within matplotlib
    fill.scale=tm_scale_intervals(
            #keep number of categories fewer. see
      #https://r-tmap.github.io/tmap/articles/basics_scales

      style="pretty",
      breaks = c(-.4,-.3,-.2,-.1,0), # you need n+1 number of breaks
            n=4,#8 categories
      values="matplotlib.plasma")
)  


qtm_rd_sc_75
```


### Target percentile: 1
```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("all_scenarios_summary_by_zcta_pt.RData")

all_scenarios_summary_by_zcta_pt_scenario_100=all_scenarios_summary_by_zcta_pt %>% 
  filter(target_percentile==1) %>% 
  #and only pick those to print
  dplyr::select(zcta,
                n_zcta_intervene,
                rd_baseline_sum,
                rd_target_sum,
                rd_diff_sum
                )

tm_100_obj= zcta_ca_geo_simplified %>% 
  left_join(all_scenarios_summary_by_zcta_pt_scenario_100, by = "zcta") 

qtm_rd_sc_100=tm_basemap("CartoDB.Positron")+
  tm_shape(tm_100_obj)+
  tm_polygons(
        lwd=0.5,
    fill = "rd_diff_sum",
    fill.legend= tm_legend("Difference in risk differences (per 100k)"),
    fill.scale=tm_scale_intervals(
      style="pretty",
      breaks = c(-.4,-.3,-.2,-.1,0), # you need n+1 number of breaks
            n=4,#8 categories
      values="matplotlib.plasma")
    ) 
qtm_rd_sc_100
```

## Comparing any two scenarios
Compare raising to the 80th percentile with the 30th percentile, for example
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 10
setwd(here("data-processed"))
load(file="scenario_compare_full_join.RData")
scenario_compare_full_join %>% 
    #a strange rounding issue. this works to get the filter to work
  filter(
    round(target_percentile_upper,2)==.8) %>% 
  filter(
    round(target_percentile_lower,2)==.3) %>% 
  ggplot(aes(x=rd_diff_diff_pt))+
  #make bins shorter to make the visualization clearer
  #.2 is too coarse;
  geom_histogram(binwidth=.05)+
#  facet_grid(rows=vars(target_percentile))+
  theme_bw(base_size = 16)+
    labs(
    y="Number of ZCTAs",
    #using \n for line break
    x="Difference in difference in risk differences\nper 100,000\n(80th: Target-Baseline)-(30th: Target-Baseline)")
```








