---
title: "Supplement"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
  bookdown::word_document2:
    toc: true
    number_sections: TRUE
---

This page is a supplement to the manuscript, entitled

**A two-stage g-computation framework for health-impact assessment using spatially varying effects**

It contains univariate and bivariate distributions of measures used in the analysis.

```{r}
#| echo: false
#| warning: false
#| message: false

#Notes:
#This document is adapted from
#climate-health/docs/wf-health-scenarios-supp.Rmd

library(here)
library(mapview)  
library(tmap)
library(leafgl)
library(tidyverse)
library(viridis)
library(viridisLite)
library(leaflet)
library(leaflet.extras)
library(tmap)
library(knitr)
library(RColorBrewer)
library(kableExtra)
library(knitr)
```


# Univariate distributions

## Rate-difference outcomes
These RDs represent the estimated number of hospitalizations attributable to each climate-related exposure per 100,000 population per event.
A positive rate difference indicates that the climate hazard (heat or wildfire) is causing hospitalizations, and a negative value implies the hazard prevents them.
Heat events are defined as two consecutive days, so estimates represent the per-heat-event burden rather than a per-day measures. 

### Heat-related hospitalizations

Histogram of Rate difference per 100,000 person-events over ZCTAs.
```{r}
#| echo: false
#| warning: false
#| message: false
options(scipen=999)
setwd(here("data-processed"))
load("outcome_heat_hi99_2_no_outliers.RData")

outcome_heat_hi99_2_no_outliers %>% 
  mutate(overall=1) %>% 
  group_by(overall) %>% 
  summarise(
    Median=median(rd_100k_quo_pt,na.rm=T),
    Mean=weighted.mean(
        x=rd_100k_quo_pt,
        w=pop
        ),
    Percentile_25th= quantile(rd_100k_quo_pt,probs=c(0.25),na.rm=T),
    Percentile_75th= quantile(rd_100k_quo_pt,probs=c(0.75),na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  knitr::kable(digits=3)


outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=rd_100k_quo_pt))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Rate difference per 100,000 person-events",
    y="Number of ZCTAs")

```



### Wildfire-related hospitalizations
Histogram of Rate difference per 100,000 person-events over ZCTAs.
```{r}
#| echo: false
#| warning: false
#| message: false
options(scipen=999)
setwd(here("data-processed"))
load("outcome_wf_no_outliers.RData")

outcome_wf_no_outliers %>% 
  mutate(overall=1) %>% 
  group_by(overall) %>% 
  summarise(
    Median=median(rd_100k_quo_pt,na.rm=T),
    Mean=weighted.mean(
        x=rd_100k_quo_pt,
        w=pop
        ),
    Percentile_25th= quantile(rd_100k_quo_pt,probs=c(0.25),na.rm=T),
    Percentile_75th= quantile(rd_100k_quo_pt,probs=c(0.75),na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  knitr::kable(digits=3)


outcome_wf_no_outliers %>% 
  ggplot(aes(x=rd_100k_quo_pt))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Rate difference per 100,000 person-events",
    y="Number of ZCTAs")

```


## Intervention measures
The intervention measures are:
* Impervious surfaces
* Tree canopy

Below are histograms of these measures over ZCTAs.

### Impervious surfaces
```{r}
#| echo: false
#| warning: false
#| message: false

outcome_wf_no_outliers %>% 
  ggplot(aes(x=imperv_prop))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Proportion Impervious Surfaces",
    y="Number of ZCTAs")
```

### Tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false

#names(outcome_wf_no_outliers)
outcome_wf_no_outliers %>% 
  ggplot(aes(x=tree_canopy_prop))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Proportion Tree Canopy",
    y="Number of ZCTAs")
```




## Covariates

### Proportion with health insurance
Distribution over ZCTAs

```{r}
#| echo: false
#| warning: false
#| message: false
#| 
#names(outcome_wf_no_outliers)
outcome_wf_no_outliers %>% 
  ggplot(aes(x=insured_prop))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Proportion with Insurance",
    y="Number of ZCTAs")
```

### Rural-urban classification

Definitions of rural-urban commuting codes:

1.  Metropolitan area core: primary flow within an urbanized area (UA)
2.  Metropolitan area high commuting: primary flow 30% or more to a UA
3.  Metropolitan area low commuting: primary flow 10% to 30% to a UA
4.  Micropolitan area core: primary flow within an Urban Cluster of 10,000 to 49,999 (large UC)
5.  Micropolitan high commuting: primary flow 30% or more to a large UC
6.  Micropolitan low commuting: primary flow 10% to 30% to a large UC
7.  Small town core: primary flow within an Urban Cluster of 2,500 to 9,999 (small UC)
8.  Small town high commuting: primary flow 30% or more to a small UC
9.  Small town low commuting: primary flow 10% to 30% to a small UC
10. Rural areas: primary flow to a tract outside a UA or UC

```{r}
#| echo: false
#| warning: false
#| message: false
setwd(here("data-processed"))
load("zcta_ca_geo_simplified.RData")
load("zcta_ruca2010.RData")
zcta_ca_geo_simplified %>% 
  left_join(zcta_ruca2010,by="zcta") %>% 
  qtm(
    basemaps = "CartoDB.Positron",
    lwd=.5,
    fill = "ruca_cat3",
#    title = "RUCA",
    fill.legend= tm_legend("RUCA category")
)
```


### Biome

```{r}
#| echo: false
#| warning: false
#| message: false
library(tmap)
setwd(here("data-input","biome-california"))
load("biomes_14_california.RData") 
biomes_14_california %>% 
  rename(biome_name=BIOME_NAME) %>% 
    qtm(
    basemaps = "CartoDB.Positron",
    lwd=.5,
    fill = "biome_name",
    fill.legend= tm_legend("Biome")
  )

```



# Bivariate Associations
In scatterplots, points are ZCTAs.

## Between RDs and intervention measures
Heat RD vs impervious surfaces

```{r}
#| echo: false
#| warning: false
#| message: false

outcome_heat_hi99_2_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=imperv_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion impervious surface",
    y="Rate difference per 100,000 person-events")

```

Heat RD vs tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false

outcome_heat_hi99_2_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=tree_canopy_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion tree canopy",
    y="Rate difference per 100,000 person-events")
```


Wildfire RD vs impervious surfaces
```{r}
#| echo: false
#| warning: false
#| message: false

outcome_wf_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=imperv_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion impervious surface",
    y="Rate difference per 100,000 person-events")

```


Wildfire RD vs tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false

outcome_wf_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=tree_canopy_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion impervious surface",
    y="Rate difference per 100,000 person-events")
```

## Between RDs and other covariates

Heat RD vs proportion insured

```{r}
#| echo: false
#| warning: false
#| message: false

outcome_heat_hi99_2_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=insured_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Rate difference per 100,000 person-events")
```


Heat RD vs rural-urban category

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=ruca_cat4, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Rural-urban category"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```


Heat RD vs biome

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=biome_name_freq, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Biome"
  )+
  #see
#  https://stackoverflow.com/questions/21878974/wrap-long-axis-labels-via-labeller-label-wrap-in-ggplot2
   scale_x_discrete(labels = scales::label_wrap(10))
```




Wildfire RD vs proportion insured

```{r}
#| echo: false
#| warning: false
#| message: false

outcome_wf_no_outliers %>%
  ggplot(aes(y=rd_100k_quo_pt,x=insured_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Rate difference per 100,000 person-events")

```



Wildfire RD vs rural-urban category

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_wf_no_outliers %>% 
  ggplot(aes(x=ruca_cat4, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Rural-urban category"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```

Wildfire RD vs biome

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_wf_no_outliers %>% 
  ggplot(aes(x=biome_name_freq, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Biome"
  )+
  #see
#  https://stackoverflow.com/questions/21878974/wrap-long-axis-labels-via-labeller-label-wrap-in-ggplot2
   scale_x_discrete(labels = scales::label_wrap(10))
```





## Between Intervention measures and other covariates

Impervious surfaces vs Proportion insured

```{r}
#| echo: false
#| warning: false
#| message: false

outcome_heat_hi99_2_no_outliers %>%
  ggplot(aes(y=imperv_prop,x=insured_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion impervious surfaces")
```


Impervious surfaces vs RUCA

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=ruca_cat4, y=imperv_prop))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Proportion impervious surfaces",
    x="Rural-urban category"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```

Impervious surfaces vs biome

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=biome_name_freq, y=imperv_prop))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Proportion impervious surfaces",
    x="Biome"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```


Tree canopy vs Proportion insured

```{r}
#| echo: false
#| warning: false
#| message: false

outcome_heat_hi99_2_no_outliers %>%
  ggplot(aes(y=tree_canopy_prop,x=insured_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion Tree canopy")
```


Tree canopy vs RUCA

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=ruca_cat4, y=tree_canopy_prop))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Proportion Tree canopy",
    x="Rural-urban category"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```

Tree canopy vs biome

```{r}
#| echo: false
#| warning: false
#| message: false
outcome_heat_hi99_2_no_outliers %>% 
  ggplot(aes(x=biome_name_freq, y=tree_canopy_prop))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Proportion Tree canopy",
    x="Biome"
  )+
   scale_x_discrete(labels = scales::label_wrap(10))
```





