---
title: "Supplement"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    toc_depth: 4
  bookdown::word_document2:
    toc: true
    number_sections: TRUE
---

```{r}
#| echo: false
#| warning: false
#| message: false

#Notes:
#This document is largely copied froim:
#climate-health-california/climate-health/docs/wf-health-scenarios-doc.Rmd

#JUly 9, 2025
#also note that not all of the modeling output looks that great in a table, so I'm going to keep it online for now and then I can adapt to word as we progress

library(here)
library(mapview)  
library(tmap)
library(leafgl)
library(tidyverse)
library(viridis)
library(viridisLite)
library(leaflet)
library(leaflet.extras)
library(tmap)
library(knitr)
library(RColorBrewer)
library(kableExtra)
library(knitr)
```

# Univariate distributions

## Outcome: Rate difference per 100,000 person-days

```{r}
#| echo: false
#| warning: false
#| message: false
options(scipen=999)
setwd(here("data-processed"))
load("wf_eff_emm_wide.RData")

wf_eff_emm_wide %>% 
  mutate(overall=1) %>% 
  group_by(overall) %>% 
  summarise(
    rd_100k_pt_med=median(rd_100k_quo_pt,na.rm=T),
    rd_100k_pt_mean_wt=weighted.mean(
        x=rd_100k_quo_pt,
        w=pop
  ),
    rd_100k_25th= quantile(rd_100k_quo_pt,probs=c(0.25),na.rm=T),
    rd_100k_75th= quantile(rd_100k_quo_pt,probs=c(0.75),na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  knitr::kable(digits=3)



wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=rd_100k_quo_pt))+
  geom_histogram()+
  theme_bw()+
  labs(x="Rate difference per 100,000 person-days")
```

## Intervention variables: Tree canopy and air conditioning

### Tree canopy 
```{r}
#| echo: false
#| warning: false
#| message: false
options(scipen=999)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>%
  mutate(overall=1) %>% 
  group_by(overall) %>%
  summarise(
    tree_canopy_prop_min=min(tree_canopy_prop,na.rm=T),
    tree_canopy_prop_max=max(tree_canopy_prop,na.rm=T),
    tree_canopy_prop_mean=mean(tree_canopy_prop,na.rm=T),
    tree_canopy_prop_med=median(tree_canopy_prop,na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  kable(digits=3)


wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=tree_canopy_prop))+
  geom_histogram()+
  theme_bw()+
  labs(x="Proportion tree canopy")

#more normal. use this instead
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=tree_canopy_prop_sqrt))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Proportion tree canopy (square root)"
  )


```

### Air conditioning
```{r}
#| echo: false
#| warning: false
#| message: false

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>%
  mutate(overall=1) %>% 
  group_by(overall) %>%
  summarise(
    ac_prop_min=min(ac_prop,na.rm=T),
    ac_prop_max=max(ac_prop,na.rm=T),
    ac_prop_mean=mean(ac_prop,na.rm=T),
    ac_prop_med=median(ac_prop,na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  kable(digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=ac_prop))+
  geom_histogram()+
  theme_bw()+
  labs(
    x="Proportion air conditioining"
  )
```

## Other socio-demographic and geographic characteristics

### Proportion above poverty
```{r}
#| echo: false
#| warning: false
#| message: false
options(scipen=999)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>%
  mutate(overall=1) %>% 
  group_by(overall) %>%
  summarise(
    above_poverty_prop_min=min(above_poverty_prop,na.rm=T),
    above_poverty_prop_max=max(above_poverty_prop,na.rm=T),
    above_poverty_prop_mean=mean(above_poverty_prop,na.rm=T),
    above_poverty_prop_med=median(above_poverty_prop,na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  kable(digits=2)


wf_eff_emm_wide %>% 
  ggplot(aes(x=above_poverty_prop))+
  geom_histogram()+
  theme_bw()+
  labs(x="Proportion above poverty")
```

### Proportion with insurance
```{r}
#| echo: false
#| warning: false
#| message: false

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>%
  mutate(overall=1) %>% 
  group_by(overall) %>%
  summarise(
    insured_prop_min=min(insured_prop,na.rm=T),
    insured_prop_max=max(insured_prop,na.rm=T),
    insured_prop_mean=mean(insured_prop,na.rm=T),
    insured_prop_med=median(insured_prop,na.rm=T)
  ) %>% 
  ungroup() %>% 
  dplyr::select(-overall) %>% 
  kable(digits=2)


wf_eff_emm_wide %>% 
  ggplot(aes(x=insured_prop))+
  geom_histogram()+
  theme_bw()+
  labs(x="Proportion with health insurance")
```

### Biome
```{r}
#| echo: false
#| warning: false
#| message: false
library(tmap)
library(sf)
setwd(here("data-input","biome-california"))
load("biomes_14_california.RData")
setwd(here("data-processed"))

biomes_14_california %>% 
  rename(biome_name=BIOME_NAME) %>% 
  mutate(
    area_m2=as.numeric(st_area(geometry)),
    area_km2=area_m2*1e-6) %>% 
  st_set_geometry(NULL) %>% 
  group_by(biome_name) %>% 
  summarise(area_km2=sum(area_km2,na.rm=T)) %>%
  arrange(desc(area_km2)) %>% 
  ungroup() %>% 
  kable(digits=0,format.args = list(big.mark = ","))
  
biomes_14_california %>% 
  rename(biome_name=BIOME_NAME) %>% 
    qtm(
    basemaps = "CartoDB.Positron",
    lwd=.5,
    fill = "biome_name",
    fill.legend= tm_legend("Biome")
  )
```

### Rural-urban classification
Definitions of rural-urban commuting codes:

1.  Metropolitan area core: primary flow within an urbanized area (UA)
2.  Metropolitan area high commuting: primary flow 30% or more to a UA
3.  Metropolitan area low commuting: primary flow 10% to 30% to a UA
4.  Micropolitan area core: primary flow within an Urban Cluster of 10,000 to 49,999 (large UC)
5.  Micropolitan high commuting: primary flow 30% or more to a large UC
6.  Micropolitan low commuting: primary flow 10% to 30% to a large UC
7.  Small town core: primary flow within an Urban Cluster of 2,500 to 9,999 (small UC)
8.  Small town high commuting: primary flow 30% or more to a small UC
9.  Small town low commuting: primary flow 10% to 30% to a small UC
10. Rural areas: primary flow to a tract outside a UA or UC
```{r}
#| echo: false
#| warning: false
#| message: false
library(tmap)
setwd(here("data-processed"))
load(file="zcta_ruca2010.RData")
load("zcta_ca_geo_simplified.RData")

zcta_ca_geo_simplified %>% 
  left_join(zcta_ruca2010,by="zcta") %>% 
  mutate(
    area_m2=as.numeric(st_area(geometry)),
    area_km2=area_m2*1e-6) %>% 
  st_set_geometry(NULL) %>% 
  group_by(ruca_cat) %>% 
  summarise(
    pop=sum(pop),
    area_km2=sum(area_km2,na.rm=T)) %>%
  ungroup() %>% 
  kable(digits=0,format.args = list(big.mark = ","))
  


zcta_ca_geo_simplified %>% 
  left_join(zcta_ruca2010,by="zcta") %>% 
  qtm(
    basemaps = "CartoDB.Positron",
    lwd=.5,
    fill = "ruca_cat",
#    title = "RUCA",
    fill.legend= tm_legend("RUCA category")
)
```

# Bivariate and stratified associations

## RD x Tree canopy x geographic measures
### RD x Tree canopy
Here are scatterplots plotting the rate difference against the (square root of) proportion tree canopy.

There is a slight negative association between tree canopy and the RD, but it is rather weak.

```{r}
#| echo: false
#| warning: false
#| message: false
library(jtools)#for summarizing glm outputs with more customization
#This look really nice, actually
#https://www.rdocumentation.org/packages/jtools/versions/2.3.0/topics/summ.glm

setwd(here("data-processed"))
load("wf_eff_emm_wide_no_outliers.RData")
glm_rd_100k_quo_pt_tree_canopy_prop_sqrt=glm(
  rd_100k_quo_pt~tree_canopy_prop_sqrt,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)

jtools::summ(
  glm_rd_100k_quo_pt_tree_canopy_prop_sqrt,
    model.fit=F, #for brevity in the output, dont' show
  digits=3)

wf_eff_emm_wide_no_outliers %>% 
  group_by(STATE) %>% 
  summarise(
    corr_spearman=cor(
      x=tree_canopy_prop_sqrt,
      y=rd_100k_quo_pt,
      method="spearman"),
    corr_pearson=cor(
      x=tree_canopy_prop_sqrt,
      y=rd_100k_quo_pt,
      method="pearson")
  ) %>% 
  dplyr::select(-STATE) %>% 
  kable(digits=2)


wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=tree_canopy_prop_sqrt, y=rd_100k_quo_pt))+
  geom_point()+
  geom_smooth(method="lm")+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion tree canopy (square root)"
  )
```


### RD x Tree canopy x biome

We stratified associations by biome, which illustrates, as above, that the tree-canopy distribution differs starkly by biome. The RD x tree canopy association appears to be negative in Mediterran Forests, Woodlands & Scrub and in Deserts & Xeric shrublands, but slightly positive in the other two.

```{r}
#| echo: false
#| warning: false
#| message: false
glm_rd_100k_quo_pt_tree_canopy_prop_sqrt_int_biome=glm(
  rd_100k_quo_pt~tree_canopy_prop_sqrt +biome_name_freq +
    biome_name_freq*tree_canopy_prop_sqrt,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)
jtools::summ(
  glm_rd_100k_quo_pt_tree_canopy_prop_sqrt_int_biome,
  model.fit=F, #for brevity in the output, dont' show
             digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=tree_canopy_prop_sqrt, y=rd_100k_quo_pt))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth(method="lm")+
  theme_bw()+
    labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion tree canopy (square root)"
    )+
  #see here:
  #https://ggplot2.tidyverse.org/reference/facet_grid.html
   theme(strip.text.y = element_text(angle = 0))
```



### RD x Tree canopy x rural-urban

We also stratified this association by urban-rural category.

The stratified scatterplot suggests that in the most urban areas and in the least urban areas, higher tree canopy is associated with a **lower** rate difference, whereas in the other two RUCA categories, the association is in the other direction: higher tree canopy is associated with a **higher** rate difference (i.e., suggesting a more harmful effect of wildfire on the outcome).

```{r}
#| echo: false
#| warning: false
#| message: false
glm_rd_100k_quo_pt_tree_canopy_prop_sqrt_int_ruca_cat=glm(
  rd_100k_quo_pt~tree_canopy_prop_sqrt + ruca_cat +ruca_cat*tree_canopy_prop_sqrt,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)
jtools::summ(
  glm_rd_100k_quo_pt_tree_canopy_prop_sqrt_int_ruca_cat,
      model.fit=F, #for brevity in the output, dont' show
  digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #remove values above the 99th percentile and below the 1st
  ggplot(aes(x=tree_canopy_prop_sqrt, y=rd_100k_quo_pt))+
  geom_point()+
  facet_grid(
    rows="ruca_cat"
  )+
  geom_smooth(method="lm")+
  theme_bw()+
    labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion tree canopy (square root)"
    )
```




## RD x A/C x geographic measures

### RD x A/C
The association with air conditioning is weakly negative.

```{r}
#| echo: false
#| warning: false
#| message: false
glm_rd_100k_quo_pt_ac_prop=glm(
  rd_100k_quo_pt~ac_prop,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)
jtools::summ(
  glm_rd_100k_quo_pt_ac_prop,
    model.fit=F, #for brevity in the output, dont' show
  digits=3)

wf_eff_emm_wide_no_outliers %>% 
  group_by(STATE) %>% 
  summarise(
    corr_spearman=cor(
      x=ac_prop,
      y=rd_100k_quo_pt,
      method="spearman",
      use="complete.obs"),
    corr_pearson=cor(
      x=ac_prop,
      y=rd_100k_quo_pt,
      method="pearson",
      use="complete.obs")
  ) %>% 
  dplyr::select(-STATE) %>% 
  kable(digits=2)



wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=ac_prop, y=rd_100k_quo_pt))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion with air conditioning"
  )
```

### RD x A/C x biome

```{r}
#| echo: false
#| warning: false
#| message: false
glm_rd_100k_quo_pt_ac_prop_int_biome=glm(
  rd_100k_quo_pt~ac_prop + biome_name_freq +biome_name_freq*ac_prop,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)
jtools::summ(
  glm_rd_100k_quo_pt_ac_prop_int_biome,
      model.fit=F, #for brevity in the output, dont' show
  digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% 
  ggplot(aes(x=ac_prop, y=rd_100k_quo_pt))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth(method="lm")+
  theme_bw()+
    labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion A/C"
    )+
  #see here:
  #https://ggplot2.tidyverse.org/reference/facet_grid.html
   theme(strip.text.y = element_text(angle = 0))
```


### RD x A/C x rural-urban

Like with tree canopy, the RD x A/C association varies by rural-urban category. It is more negative in more populous areas, and positive–suggesting air conditioning is associated with a **higher** difference effect of wildfire on acute-care utilization in less populous areas.

```{r}
#| echo: false
#| warning: false
#| message: false
glm_rd_100k_quo_pt_ac_prop_int_ruca_cat=glm(
  rd_100k_quo_pt~ac_prop + ruca_cat +ruca_cat*ac_prop,
  family = gaussian,
  na.action = "na.exclude", 
  weights = rd_quo_weight,
  data = wf_eff_emm_wide_no_outliers
)
jtools::summ(
  glm_rd_100k_quo_pt_ac_prop_int_ruca_cat,
      model.fit=F, #for brevity in the output, dont' show
  digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #remove values above the 99th percentile and below the 1st
  ggplot(aes(x=ac_prop, y=rd_100k_quo_pt))+
  geom_point()+
  facet_grid(
    rows="ruca_cat"
  )+
  geom_smooth(method="lm")+
  theme_bw()+
    labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion with air conditioning"
    )
```



## RD x socio-demographic and geographic measures

### RD x proportion above poverty

```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=rd_100k_quo_pt))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion above poverty"
  )
```

### RD x proportion insured
The association between the RD and proportion with insurance is stronger than that between proportion above poverty.
```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=rd_100k_quo_pt))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Proportion above poverty"
  )
```



### RD x biome

The RD distribution is in the positive direction in **Mediterranean Forests, Woodlands & Scrub** and more negative in all the others.

```{r}
#| echo: false
#| warning: false
#| message: false

wf_eff_emm_wide %>% 
  group_by(biome_name_freq) %>% 
  summarise(
        n_zcta=n(),
    rd_100k_pt_med=median(rd_100k_quo_pt,na.rm=T),
    rd_100k_pt_mean_wt=weighted.mean(
        x=rd_100k_quo_pt,
        w=pop,
        na.rm=T
  )
  ) %>% 
  knitr::kable(digits=3)

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=biome_name_freq, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Biome"
  )+
  #see
#  https://stackoverflow.com/questions/21878974/wrap-long-axis-labels-via-labeller-label-wrap-in-ggplot2
   scale_x_discrete(labels = scales::label_wrap(10))
```

### RD x urban-rural category

The distribution is more negative in more rural areas, suggesting in those areas, wildfires had a **preventive** effect on healthcare utilization

```{r}
#| echo: false
#| warning: false
#| message: false

wf_eff_emm_wide %>% 
  group_by(ruca_cat) %>% 
  summarise(
        n_zcta=n(),
    rd_100k_pt_med=median(rd_100k_quo_pt,na.rm=T),
    rd_100k_pt_mean_wt=weighted.mean(
        x=rd_100k_quo_pt,
        w=pop,
        na.rm=T
  )) %>% 
  knitr::kable(digits=3)



wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=ruca_cat, y=rd_100k_quo_pt))+
  geom_boxplot()+
  theme_bw()+
  labs(
    y="Rate difference per 100,000 person-days",
    x="Rural-urban category"
  )+
  #see
#  https://stackoverflow.com/questions/21878974/wrap-long-axis-labels-via-labeller-label-wrap-in-ggplot2
   scale_x_discrete(labels = scales::label_wrap(10))
```

## Tree canopy and AC x socio-demographic measures
### Proportion above poverty x tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=tree_canopy_prop_sqrt))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion above poverty",
    y="Proportion tree canopy (square root)"
  )

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=tree_canopy_prop_sqrt))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion above poverty",
    y="Proportion tree canopy (square root)"
  )
```

### Proportion insured x tree canopy
```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=insured_prop, y=tree_canopy_prop_sqrt))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion tree canopy (square root)"
  )

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=insured_prop, y=tree_canopy_prop_sqrt))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion tree canopy (square root)"
  )
```

### Proportion above poverty x A/C
```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=ac_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion above poverty",
    y="Proportion with A/C"
  )

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=above_poverty_prop, y=ac_prop))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion above poverty",
    y="Proportion with A/C"
  )
```

### Proportion insured x A/C
```{r}
#| echo: false
#| warning: false
#| message: false
wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=insured_prop, y=ac_prop))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion with A/C"
  )

wf_eff_emm_wide %>% 
  filter(rd_quo_outlier==0) %>% #
  ggplot(aes(x=insured_prop, y=ac_prop))+
  geom_point()+
  facet_grid(
    rows="biome_name_freq",
    labeller = label_wrap_gen(20)#wrap text!
  )+
  geom_smooth()+
  theme_bw()+
  labs(
    x="Proportion insured",
    y="Proportion with A/C"
  )
```




